---
title: "Cultural Cognition, Environmental Orientation, and Pro-Environmental Behavior"
bibliography: references.bib
csl: american-political-science-association.csl
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  html: default
  pdf:
    include-in-header:
      text: |
        \usepackage{setspace}
        \usepackage{dcolumn}
        \usepackage{float}
        \usepackage{amsmath}
        \addtokomafont{disposition}{\rmfamily}
  docx:
    reference-doc: sustainability-template.dot
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Clean, fit, and build all tables/figures (no side effects). Robust to cwd.
source(if (file.exists("scripts/analysis-prep.R")) "scripts/analysis-prep.R" else "../scripts/analysis-prep.R")
```

# Introduction

# Belief Systems and Pro-Environmental Behavior 


## Cultural Cognition


## New Ecological Paradigm


## Connectedness to Nature 


## Pro-Environmental Behaviors 


# Hypotheses


# Data and Methods 

## Measures 

<!-- Table 1 -->

```{r}
#| label: tbl-behaviors
#| tbl-cap: "Self-reported pro-environmental behaviors by type (public/private), with means and standard deviations"
knitr::kable(tbl1, booktabs = TRUE, align = c("l", "l", "r", "r"))
```


<!-- Table 2 -->

```{r}
#| label: tbl-measurement
#| tbl-cap: "Measurement items and construct reliability (Cronbach's α, composite reliability, average variance extracted)"
knitr::kable(tbl2, booktabs = TRUE, align = c("l", "l", "r", "r", "r", "r", "r"))
```


# Results

<!-- Combined serial model: cultural cognition -> NEP -> CNS -> behavior.
     The CNS-only model, NEP-only model, their fit table, and the separate
     mediation analysis now live in cc-behave-supplemental.qmd.
     Reference these in your prose with @fig-serial / @tbl-serial-fit /
     @tbl-serial-effects. -->

<!-- Figure: full combined model path diagram -->

```{r}
#| label: fig-serial
#| fig-cap: "Full combined structural model: cultural cognition → New Ecological Paradigm → Connectedness to Nature → pro-environmental behavior, with direct paths. Standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6.5
draw_sem(fitFull)
```

<!-- Table 5: serial vs full fit comparison -->

```{r}
#| label: tbl-serial-fit
#| tbl-cap: "Fit comparison of the strict serial chain versus the full combined model with direct paths added"
knitr::kable(tbl5, booktabs = TRUE, align = c("l", rep("r", 9)))
```

`r lrt_text`

<!-- Table 6: effect decomposition in the full model -->

```{r}
#| label: tbl-serial-effects
#| tbl-cap: "Standardized decomposition of cultural cognition effects on public and private behavior in the full combined model († p<0.10, * p<0.05, ** p<0.01, *** p<0.001)"
knitr::kable(tbl6, booktabs = TRUE, align = c("l", "l", rep("r", 4)))
```

# Discussion


# Conclusion 

\newpage

# Appendix

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\newpage

# References

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\singlespace

\noindent