---
title: "Supplemental Materials: Cultural Cognition, Environmental Orientation, and Pro-Environmental Behavior"
bibliography: references.bib
csl: american-political-science-association.csl
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  html: default
  pdf:
    include-in-header:
      text: |
        \usepackage{setspace}
        \usepackage{dcolumn}
        \usepackage{float}
        \usepackage{amsmath}
        \addtokomafont{disposition}{\rmfamily}
  docx:
    reference-doc: sustainability-template.dot
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Shares the same analysis engine as the main paper (no side effects). Robust to cwd.
source(if (file.exists("scripts/analysis-prep.R")) "scripts/analysis-prep.R" else "../scripts/analysis-prep.R")
```

These supplemental materials report the single-orientation structural equation models — cultural cognition predicting the Connectedness to Nature Scale (CNS) and the New Ecological Paradigm (NEP) separately — the associated mediation analysis, and the full combined model estimated with the four separate worldview measures (hierarchy, egalitarianism, individualism, communitarianism). The main text reports the combined model with the four worldview measures collapsed into two composite groups, Hierarch-Individualist (HIER-INDIV) and Egalitarian-Communitarian (EGAL-COMM); the four-measure version here is the specification the composite-group model replaces.

# Single-orientation structural models

<!-- Figure S1: CNS model -->

```{r}
#| label: fig-cns
#| fig-cap: "Measurement and structural model of cultural cognition, the Connectedness to Nature Scale, and public and private pro-environmental behaviors. Controlling for age, gender, race, education, and household income (paths not shown). Standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6
draw_sem(fitCNS)
```

<!-- Figure S2: NEP model -->

```{r}
#| label: fig-nep
#| fig-cap: "Measurement and structural model of cultural cognition, the New Ecological Paradigm, and public and private pro-environmental behaviors. Controlling for age, gender, race, education, and household income (paths not shown). Standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6
draw_sem(fitNEP)
```

<!-- Table S1: CNS/NEP model fit -->

```{r}
#| label: tbl-fit
#| tbl-cap: "Model fit statistics for the CNS and NEP single-orientation structural equation models. Both models control for age, gender, race, education, and household income."
knitr::kable(tbl3, booktabs = TRUE, align = c("l", rep("r", 8)))
```

# Full combined model with the four separate worldview measures

This section reports the full combined model (cultural worldviews → NEP → CNS → behavior, with direct paths) estimated with the four separate worldview measures — hierarchy, egalitarianism, individualism, and communitarianism — as latent predictors. The main text reports the same model with these four measures collapsed into the two composite groups described above; this specification is the one the composite-group model replaces.

<!-- Figure S3: structural model only (measurement model omitted) -->

```{r}
#| label: fig-structural-4wv
#| fig-cap: "Structural model of the full combined SEM with the four separate worldview measures: cultural cognition → New Ecological Paradigm → Connectedness to Nature → public and private pro-environmental behavior, with direct paths. Controlling for age, gender, race, education, and household income (paths not shown). The measurement (indicator) portion is omitted for clarity; standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6.5
draw_sem_struct(fitFull)
```

<!-- Table S2: serial vs full fit comparison, four worldview measures -->

```{r}
#| label: tbl-serial-fit-4wv
#| tbl-cap: "Fit comparison of the strict serial chain versus the full combined model with direct paths added, four separate worldview measures. Both models control for age, gender, race, education, and household income."
knitr::kable(tbl5, booktabs = TRUE, align = c("l", rep("r", 9)))
```

`r lrt_text`

<!-- Table S3: effect decomposition in the full model, four worldview measures -->

```{r}
#| label: tbl-serial-effects-4wv
#| tbl-cap: "Standardized decomposition of cultural cognition effects on public and private behavior in the full combined model, four separate worldview measures, controlling for age, gender, race, education, and household income († p<0.10, * p<0.05, ** p<0.01, *** p<0.001)"
knitr::kable(tbl6, booktabs = TRUE, align = c("l", "l", rep("r", 4)))
```

# Full combined model with measurement detail

Figure S4 shows the same four-worldview-measure full combined model with the full measurement model — the latent constructs and their indicators — included.

<!-- Figure S4: full combined model, measurement + structure -->

```{r}
#| label: fig-serial
#| fig-cap: "Full combined model including the measurement model (latent constructs and their indicators) and the structural paths: cultural cognition → New Ecological Paradigm → Connectedness to Nature → public and private pro-environmental behavior, with direct paths. Controlling for age, gender, race, education, and household income (paths not shown). Standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6.5
draw_sem(fitFull)
```

# Mediation analysis

<!-- Table S2: mediation -->

```{r}
#| label: tbl-mediation
#| tbl-cap: "Mediation analysis: standardized direct, indirect, and total effects of cultural cognition on public and private pro-environmental behavior, controlling for age, gender, race, education, and household income († p<0.10, * p<0.05, ** p<0.01, *** p<0.001)"
knitr::kable(tbl4, booktabs = TRUE, align = c("l", "l", "r", "r", "r"))
```

# Demographic control coefficients

Every model above and in the main text controls for age, gender (male), race (white), education, and household income, but these coefficients are omitted from the path diagrams and decomposition tables for readability. Table S6 reports the standardized control coefficients from the full combined model reported in the main text (HIER-INDIV and EGAL-COMM in place of the four separate worldview measures; main-text Table 6).

<!-- Table S6: demographic control coefficients, main-text full model -->

```{r}
#| label: tbl-hiec-controls
#| tbl-cap: "Standardized demographic control coefficients from the full combined model reported in the main text (HIER-INDIV/EGAL-COMM), by outcome († p<0.10, * p<0.05, ** p<0.01, *** p<0.001)"
knitr::kable(tbl_hiec_controls, booktabs = TRUE, align = c("l", "r", "r", "r", "r"))
```

\newpage

# References

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\singlespace

\noindent