---
title: "Supplemental Materials: Cultural Cognition, Environmental Orientation, and Pro-Environmental Behavior"
bibliography: references.bib
csl: american-political-science-association.csl
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  html: default
  pdf:
    include-in-header:
      text: |
        \usepackage{setspace}
        \usepackage{dcolumn}
        \usepackage{float}
        \usepackage{amsmath}
        \addtokomafont{disposition}{\rmfamily}
  docx:
    reference-doc: sustainability-template.dot
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Shares the same analysis engine as the main paper (no side effects). Robust to cwd.
source(if (file.exists("scripts/analysis-prep.R")) "scripts/analysis-prep.R" else "../scripts/analysis-prep.R")
```

These supplemental materials report the single-orientation structural
equation models — cultural cognition predicting the Connectedness to
Nature Scale (CNS) and the New Ecological Paradigm (NEP) separately — and
the associated mediation analysis. The combined model that integrates both
orientations is reported in the main text.

# Single-orientation structural models

<!-- Figure S1: CNS model -->

```{r}
#| label: fig-cns
#| fig-cap: "Measurement and structural model of cultural cognition, the Connectedness to Nature Scale, and public and private pro-environmental behaviors. Standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6
draw_sem(fitCNS)
```

<!-- Figure S2: NEP model -->

```{r}
#| label: fig-nep
#| fig-cap: "Measurement and structural model of cultural cognition, the New Ecological Paradigm, and public and private pro-environmental behaviors. Standardized coefficients shown; * p<0.05, ** p<0.01, *** p<0.001."
#| fig-width: 9
#| fig-height: 6
draw_sem(fitNEP)
```

<!-- Table S1: CNS/NEP model fit -->

```{r}
#| label: tbl-fit
#| tbl-cap: "Model fit statistics for the CNS and NEP single-orientation structural equation models"
knitr::kable(tbl3, booktabs = TRUE, align = c("l", rep("r", 8)))
```

# Mediation analysis

<!-- Table S2: mediation -->

```{r}
#| label: tbl-mediation
#| tbl-cap: "Mediation analysis: standardized direct, indirect, and total effects of cultural cognition on public and private pro-environmental behavior († p<0.10, * p<0.05, ** p<0.01, *** p<0.001)"
knitr::kable(tbl4, booktabs = TRUE, align = c("l", "l", "r", "r", "r"))
```

\newpage

# References

\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\singlespace

\noindent
