---
title: "Behaving Sustainably: Cultural Cognition, Environmental Orientation, and Pro-Environmental Behavior"
bibliography: references.bib
csl: american-political-science-association.csl
fig-cap-location: top
fontfamily: ebgaramond
link-bibliography: false
format:
  html:
    author:
      - name: Matthew C. Nowlin
        email: matthew.nowlin@uta.edu
        corresponding: true
        affiliations:
          - name: University of Texas at Arlington
            department: Department of Political Science
      - name: Aaron Sparks
        affiliations: 
          - name: University of North Carolina – Chapel Hill
            deparment: School of Government
    template-partials:
      - title-metadata.html
  pdf:
    author:
      - name: "Matthew C. Nowlin^[Associate Professor, Department of Political Science, University of Texas at Arlington. Email: matthew.nowlin@uta.edu]"
      - name: "Aaron Sparks^[Adjunct Instructor, School of Government, University of North Carolina – Chapel Hill]"
    include-in-header:
      text: |
        \usepackage{setspace}
        \usepackage{dcolumn}
        \usepackage{float}
        \usepackage{amsmath}
        \addtokomafont{disposition}{\rmfamily}
        \normalem
  docx:
    author:
      - name: "Matthew C. Nowlin^[Associate Professor, Department of Political Science, University of Texas at Arlington. Email: matthew.nowlin@uta.edu]"
      - name: "Aaron Sparks^[Adjunct Instructor, School of Government, University of North Carolina – Chapel Hill]"
    reference-doc: custom-reference-doc.docx
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Clean, fit, and build all tables/figures (no side effects). Robust to cwd.
source(if (file.exists("scripts/analysis-prep.R")) "scripts/analysis-prep.R" else "../scripts/analysis-prep.R")
```

**Abstract**: Scholars has long noted an "attitude-behavior gap" which reflects an incomplete understanding of how values and beliefs translate into pro-environmental action. This study develops and tests a conceptual model in which cultural worldviews along grid (hierarchy-egalitarianism) and group (individualism-communitarianism) dimensions shape environmental orientations, which in turn drive pro-environmental behavior. Environmental orientation is treated as having a cognitive component, measured by the New Ecological Paradigm (NEP), and an affective component, measured by the Connectedness to Nature Scale (CNS). Using survey data from 501 US adults collected online in July 2017, we estimate a Structural Equation Model (SEM) to test direct, indirect, and mediated relationships among cultural worldviews, the NEP, the CNS, and a 12-item self-reported behavior scale. Results showed hierarchical individualists held weaker ecological worldviews, while egalitarian communitarians held stronger ecological worldviews and felt more connected to nature. Both the NEP and CNS were positively associated with behavior, with the CNS showing the stronger relationship. Cultural worldviews had no significant direct effect on behavior, rather their influence operated entirely through environmental orientations. These findings indicate that cultural worldviews shape pro-environmental behavior indirectly, mediated by cognitive and especially affective environmental orientations, helping to explain part of the attitude-behavior gap.

**Keywords**: Pro-environmental behaviors; cultural cognition; ecological worldview; connected to nature  

# Introduction

Achieving environmentally sustainable outcomes likely requires some behavioral changes at the individual, community, and societal levels. However, not all individuals are equally likely to engage in pro-environmental behaviors, and understanding the psychological and cultural factors that drive these behaviors is important for enhancing sustainability. Environmental behaviors are widely studied [@zhuBibliometricVisualAnalysis2021], and scholars of environmental psychology often lament the so-called ‘attitude-behavior’ gap wherein environmental attitudes do not explain very much of the variance in environmental behavior [@sternNewEnvironmentalTheories2000; @kollmussMindGapWhy2002; @giffordPersonalSocialFactors2014]. If people care strongly about the environment, why do they not act in a more sustainable way? The attitude-behavior gap is not because people are hypocrites but because researchers do not yet have a complete attitude-behavior model [@kaiserOffsettingBehavioralCosts2021].

To understand how environmental attitudes influence behaviors, scholars often use approaches such as the Value-Belief-Norm (VBN) framework, where higher-order values shape second-order beliefs that, in turn, influence behaviors [@canlasMoreTwentyYears2022; @sternNewEnvironmentalTheories2000; @sternValueBeliefNormTheorySupport1999; @vysevkovaExplainingSupportEnvironmental2026]. Higher-order values form a cognitive foundation that influences how information about the world is processed and understood across multiple issues [@ripbergerCulturalTheoryMeasurement2014]. When considering environmental behaviors, the second-order values involve environmental cognition, such as views about the relationship of humans to the natural world and how information about the environment is understood [@sternValueBasisEnvironmental1994; @henryUnderstandingEnvironmentalCognition2012]. In addition to environmental cognition, individuals also have affective feelings about the environment [@gkargkavouziMultidimensionalMeasureEnvironmental2019; @mayerConnectednessNatureScale2004]. Together, environmental cognition and affective feelings form an environmental orientation that shapes environmental behaviors. However, to develop a complete attitude-behavior model, it is important to understand the various links in the chain, including the higher-order values and cognition that inform environmental orientations. Of note, the cultural worldviews measured by cultural cognition have been shown to shape environmental orientations [@nowlinEnvironmentalPolicymakingEra2019].

Cultural cognition was developed as a tool to measure how cultural values shape beliefs about risk [e.g., @kahanCultureIdentityProtectiveCognition2007]. It is based on cultural theory and its conceptualization of worldviews [@kahanCulturalCognitionConception2012; @douglasRiskCultureEssay1982]. Cultural theory addresses how cultural values influence people’s perceptions of and responses to risk. By measuring an individual’s cultural values, we may be able to predict how they will perceive and respond to risks associated with certain activities or products. We are interested in how cultural cognition shapes views toward environmental issues and how these views shape behavior. Cultural worldviews and their subsequent risk perceptions are likely strong drivers of environmental orientation because environmental problems typically involve elements of risk or harm to the environment. 

The cultural worldviews associated with cultural cognition are measured in two dimensions, a grid dimension (egalitarianism-hierarchy) and a group dimension (communitarianism-individualism). In general, those who are more egalitarian and communitarian are more concerned about environmental risks than those who are more hierarchical and individualist. Therefore, it is expected that individuals high in egalitarianism and communitarianism should engage in pro-environmental behaviors to mitigate risks to the natural environment. However, previous research has shown only a weak correlation between egalitarianism, communitarianism, and pro-environmental behavior [@sparksMeasuringEnvironmentalValues2021]. Therefore, cultural cognition shapes how individuals think and process information about issues but may not directly influence behaviors.

In this paper, we develop a conceptual framework that accounts for both the cognitive and affective influences on pro-environmental behavior. Specifically, we posit that higher-order beliefs, drawn from cultural cognition [@kahanCulturalCognitionConception2012; @kahanCulturalCognitionPublic2006; @kahanPolarizingImpactScience2012], inform environmental orientations [@henryUnderstandingEnvironmentalCognition2012], which in turn influence both public and private behavior. Drawing from @henryUnderstandingEnvironmentalCognition2012, we use the New Ecological Paradigm (NEP) as our measure of environmental cognition and the Connectedness to Nature Scale (CNS) as the affective element of environmental orientation. Previous work failed to find that cultural cognition influences environmental behavior; however, it has been found to influence environmental cognition using the NEP [@nowlinEnvironmentalPolicymakingEra2019]. Therefore, we hypothesize that environmental orientations mediate the relationship between cultural cognition and behavior.

Overall, we found that egalitarian communitarians hold an ecological worldview and feel more connected to nature. However, hierarchical individualists are less concerned about the environment than those with other worldviews. Additionally, we find that ecological worldviews are associated with feeling connected to nature, and that both worldviews and affective nature connection are associated with an increase in self-reported pro-environmental behaviors. Finally, we find that environmental orientations mediate the relationship between cultural worldviews and behaviors, meaning that cultural worldviews play an important, but indirect, role in environmental behaviors. 


# Cognition, Environmental Orientations, and Pro-Environmental Behavior

The process through which values and beliefs impact behavior is often understood as a process where higher-order, antecedent values influence second-order beliefs, which in turn influence behavior [@stegGeneralAntecedentsPersonal2011; @dietzEnvironmentalValues2005]. One of the leading frameworks that adopts this approach is the Value-Belief-Norm (VBN) framework [@sternValueBeliefNormTheorySupport1999; @sternNewEnvironmentalTheories2000]. In brief, the VBN posits that underlying values inform broad environmental orientations that activate norms that, in turn, support pro-environmental behavior. VBN involves a sequential chain that goes from values to beliefs to behavior. Therefore, values inform attitudes, which in turn inform behavior.

As noted, the VBN framework moves from values to beliefs to norms to pro-environmental behaviors. The values that have been identified and applied include biospheric, altruistic, and egoistic values [@canlasMoreTwentyYears2022; @sternNewEnvironmentalTheories2000]. Biospheric values are associated with concern for the environment, altruistic values involve concern for others, and egoistic values shape concern for self [@fornaraExtendedValueBeliefNormTheory2020]. In general, biospheric and altruistic values are associated with an ecological worldview, whereas egoistic values are not. Beliefs are grounded in ecological worldviews, which are typically operationalized using the New Ecological Paradigm (NEP) measure [@chenExaminationValuebeliefnormTheory2015; @fornaraExtendedValueBeliefNormTheory2020; @slimakPersonalValuesBeliefs2006; @sternValueBeliefNormTheorySupport1999]. Additional beliefs include awareness of consequences (AC) and ascription of responsibility (AR), which are active norms involving a sense of obligation to act [@sternNewEnvironmentalTheories2000]. The behaviors identified include private consumption-based behaviors, public-facing activism, and support for environmental policies [@sternValueBeliefNormTheorySupport1999]. VBN is widely used [@canlasMoreTwentyYears2022; @vysevkovaExplainingSupportEnvironmental2026] across multiple contexts, including climate mitigation behaviors in Taiwan [@chenExaminationValuebeliefnormTheory2015], energy conservation [@yeboahExplainingEnergyConservation2016], biodiversity in Europe [@fornaraExtendedValueBeliefNormTheory2020; @riepeValuesBeliefsNorms2021], and perceptions of environmental risk [@slimakPersonalValuesBeliefs2006].

A similar set of assumptions inform an understanding of belief systems developed by policy scholars using the Advocacy Coalition Framework (ACF) [@nohrstedtAdvocacyCoalitionFramework2023; @sabatierPolicyChangeLearning1993; @henryUnderstandingEnvironmentalCognition2012]. With this approach, belief systems are interrelated beliefs made up of three different types, including deep core, policy core, and secondary beliefs. Deep core beliefs are the broadest in scope and are fundamental beliefs that inform attitudes toward multiple issues. They tend to be stable, play a role in how issues are understood, and influence reasoning about issues [@kahanPolarizingImpactScience2012; @nohrstedtAdvocacyCoalitionFramework2023]. Biospheric, altruistic, and egoistic values, as well as political ideology, can be seen as examples of deep core beliefs. Next, are policy core beliefs, which are beliefs about particular issues. Environmental orientations would be akin to policy core beliefs. Finally, secondary aspects are the narrowest in scope and include beliefs about how to best achieve the goals outlined in policy core beliefs. We aim to build on previous work that noted a connection between the VBN and the ACF [@henryUnderstandingEnvironmentalCognition2012].

In developing our approach, we draw from both VBN and the ACF by positing a sequential chain and a hierarchical ordering of values, beliefs, and attitudes that lead from abstract values to concrete actions. Specifically, we note that deep core beliefs inform environmental orientations that, in turn, make pro-environmental behaviors more (or less) likely. In addition, we separate beliefs that shape *cognition* from the affective and identity elements of environmental orientations. We note that the role of cognition has not been well developed in approaches to understanding environmental behaviors, despite the importance that the way worldviews shape information processing has on midstream beliefs and downstream behaviors [@liobikieneImportanceEnvironmentalKnowledge2019; @ripbergerCulturalTheoryMeasurement2014]. Our measure of deep core beliefs is cultural cognition, which, as the name implies, involves deeply rooted values and beliefs that influence cognition. For environmental orientations, we draw on environmental cognition, which is "the way individuals structure their thinking about environmental issues and associated political actions" [@henryUnderstandingEnvironmentalCognition2012, 238]. We argue that environmental orientations contain both cognitive and affective elements. Deep core beliefs influence both the cognitive and affective aspects of environmental orientations. Finally, deep core beliefs and environmental orientations influence pro-environmental behaviors; however, it is not clear whether deep core beliefs influence behaviors directly [@markleUnderstandingProEnvironmentalBehavior2019] or indirectly [@sparksMeasuringEnvironmentalValues2021]. However, the affective element of environmental orientations tends to have the most direct influence on behaviors [@sparksMeasuringEnvironmentalValues2021; @mackayPeopleWhoFeel2019].

## Deep Core Beliefs: Cultural Cognition

For our measure of deep core beliefs we use cultural cognition. Cultural cognition is derived from the cultural theory of risk developed by Mary Douglas and Aaron Wildavsky [@douglasRiskCultureEssay1982; @thompsonCulturalTheory1990; @kahanCulturalCognitionConception2012]. Cultural theory (CT) has been identified as a leading way to conceptualize and measure deep core beliefs [@ripbergerCulturalTheoryMeasurement2014]. In brief, CT posits that social structures and relationships are represented by two dimensions, including a grid dimension and a group dimension. Four competing cultural worldviews are possible based on their positioning in the grid and group dimensional space. The grid dimension represents preferred social structures regarding the roles that one plays in society and the rules by which one must abide. The group dimension represents the importance of cohesive social units and the well-being of the broader community. The original formulation of CT produces four cultural worldviews based on the relative positions on the grid and group dimensions, including hierarchy (high grid, high group), egalitarianism (low grid, high group), individualism (low grid, and low group), and fatalism (high grid, low group).

Cultural worldviews provide a foundational set of deep core beliefs that inform beliefs about specific policy issues (i.e., policy core beliefs), societal controversies, and risks, among other beliefs [@nowlinEnvironmentalPolicymakingEra2019; @ripbergerCulturalTheoryMeasurement2014; @jenkins-smithBeliefSystemContinuity2014; @sotirovCognitiveTheoryShifting2016]. In the context of cultural worldviews, perceptions of risk are based on the degree to which they undermine or threaten the preferred social order within each worldview. Several studies have shown a connection between cultural worldviews and risk perceptions associated with a host of environmental issues [see @xueCulturalWorldviewsEnvironmental2014], including fracking [@boudetFrackingControversyCommunication2014], nuclear waste [@jenkins-smithIdeologyCultureRisk1994], renewable energy [@tumlisonWhatExplainsLocal2018], geoengineering [@kahanGeoengineeringClimateChange2015], natural resource management [@quanCognitionConservationApplying2025], high-voltage power lines [@moyerCulturalPredispositionsSpecific2019], and climate change [@jonesLeadingWayCompromise2011; @kahanPolarizingImpactScience2012; @nowlinCulturalTheoryPartisanship2020; @nowlinWhoShouldMore2022], among others.

Cultural cognition is based on CT and its cultural worldviews but makes two important distinctions. The first is its focus on the social and psychological mechanisms that underlie how worldviews influence opinions on contested political and policy issues. Cultural commitments occur prior to facts, "in the cognitive sense that what citizens believe about the empirical consequences of \[controversial\] policies *derives* from their cultural worldviews" [@kahanCulturalCognitionPublic2006, 148]. Individuals rely on five mechanisms to ensure that their positions on issues are congruent with their cultural worldviews. These mechanisms include: 1) *cultural identity-protective cognition*, where individuals conform their beliefs to those in their "group" that share cultural values; 2) *culturally biased assimilation and cultural polarization*, which involves selective attention to worldview-confirming information and balanced information that tends to drive people of opposing worldviews further apart rather than together; 3) *cultural availability*, which occurs as worldview-confirming events and experts are more readily noticed, credited, and recalled; 4) *cultural credibility heuristic*, when people trust sources whose apparent values match their own; and 5) *cultural-identity affirmation*, which has been shown when framing information alongside an identity-affirming rather than identity-threatening policy solution makes people more receptive to the underlying facts [@kahanCulturalCognitionConception2012]. Together, these mechanisms account for the cognitive aspect of cultural cognition.

The second distinction between cultural cognition and CT is how cultural worldviews are measured empirically. As noted, the CT worldviews are based on the placement on the grid and group dimensions, and worldviews are a result of placement on both. However, cultural cognition conceives grid and group as two orthogonal dimensions, with worldviews identified as being high or low on each dimension rather than a combination of the dimensions. Those high on the grid dimension are called hierarchical, as they prefer a society with well-defined roles and clear lines of authority, whereas those low on the grid dimension are termed egalitarian, as they value a social order based on an equal sharing of power and resources. Communitarianism is the worldview that is high on the group dimension and represents a preference for a collectivist society. Finally, individualism is low on the group dimension, preferring increased individual autonomy and responsibility [@kahanCulturalCognitionConception2012; @kahanCulturalCognitionPublic2006; @kahanPolarizingImpactScience2012].

Studies using cultural cognition often compare the risk perceptions of hierarchical individualists, those high on the grid dimension and low on the group dimension, to their reverse, egalitarian communitarians who are low on grid and high on the group dimension. The differences in risk perception between these pairings are significant for risks related to the environment, abortion, and guns [@kahanCultureIdentityProtectiveCognition2007]. For environmental risks specifically, communitarianism, with its concern about collective well-being and egalitarianism, with its concern about the negative impacts of concentrated power tend to see threats to the natural world as a large risk. Hierarchs believe that natural resources can be well managed by experts and individualism supports the use of natural resources in private enterprise, and therefore hierarchy and individualism tend not to be concerned about environment harm. Studies have shown the differences between hierarchical individualists and egalitarian communitarians on climate change [@kahanPolarizingImpactScience2012] and water supply risk [@overdevestUsingCulturalCognition2013]. In terms of the influence of cultural cognition and pro-environmental behavior, the results are somewhat mixed. One study found that hierarchy and individualism predicted fewer self-reported pro-environment behaviors [@markleUnderstandingProEnvironmentalBehavior2019], whereas another study found no connection between cultural cognition and behavior [@sparksMeasuringEnvironmentalValues2021].

As a measure of deep core beliefs, cultural cognition will likely have an indirect association with environmental behaviors, meaning that the relationship will be mediated by environmental orientations. Indeed, cultural cognition worldviews, specifically egalitarianism and communitarianism are associated with pro-environmental orientation, specifically the New Ecological Paradigm (NEP) [@nowlinEnvironmentalPolicymakingEra2019; @overdevestUsingCulturalCognition2013].

## New Ecological Paradigm

Moving from deep core beliefs to environmental orientations, the first measure of environmental orientations is the New Ecological Paradigm (NEP). The NEP, developed in 1978 [@dunlapNewEnvironmentalParadigm1978], is often used in the VBN framework as a measure of environmental beliefs [@canlasMoreTwentyYears2022], and has been widely adopted to explain environmental behavior [e.g., @sternNewEnvironmentalTheories2000]. Its original development focused on explaining the values associated with environmentalism as an emergent environmental paradigm that challenged the dominant social paradigm. Central to the NEP is the concern about the environment and beliefs about the proper relationship between humans and the natural world [@sternNewEcologicalParadigm1995]. The NEP was updated in 2000, and the new measure comprises five aspects which are “the reality of the limits to growth, anti-anthropocentrism, the fragility of earth’s balance, rejection of exemptionalism, and the possibility of an ecocrisis” [@dunlapNewTrendsMeasuring2000, 432].

In our approach, we view the NEP as an ecological worldview that encompasses values associated with the environment, and like cultural cognition, works to ensure that an individual's beliefs about the environment remain congruent with their ecological worldview. As a worldview, the NEP measures the environmental cognition that shapes how the natural world is seen [@xiaoEcologicalWorldviewCentral2019; @ballewSystemsThinkingPathway2019; @henryUnderstandingEnvironmentalCognition2012]. Similar to how the NEP has been used in the VBN framework, we consider it an environmental orientation that is situated between values and behaviors; however, we argue that, as a worldview, it is a cognitive component of environmental orientations. As a cognition, it influences thinking about environmental issues, shapes other aspects of environmental orientations, and is associated with pro-environmental behaviors [@derdowskiNewEcologicalParadigm2020].

There are some significant critiques of the NEP. One study applying scale analysis of the NEP based on a random sample of Norwegians found that the scale is not unidimensional, as is often claimed [@grendstadNewEcologicalParadigm1999]. In another study, @sparksMeasuringProenvironmentalOrientation2022 criticized the NEP for failing to adequately identify political conservatives who also hold a pro-environment orientation. Most important for our purposes is a study that found the NEP to be a relatively weak predictor of self-reported pro-environmental behavior, accounting for 12% of the variation when compared in the same sample to the Connectedness to Nature Scale and environmentalist identity [@sparksMeasuringEnvironmentalValues2021]. This may be because endorsement of the NEP has decreased over time [@louMetaanalysisTemporalShifts2022], or it may be that feeling close and connected to nature is more proximate to behavior than more abstract, cognitive worldviews, such as cultural cognition and the NEP.

## Connectedness to Nature

Our second instrument for measuring environmental orientation is the Connectedness to Nature Scale (CNS) [@mayerConnectednessNatureScale2004]. The CNS seeks to measure an individual’s emotional connection to nature rather than a cognitive environmental worldview. The CNS is based on the notion that to address environmental problems, people must feel emotionally connected to the non-human community on which they are dependent. The idea is that the current ecological crises are a direct result of humanity losing touch with the broader community in which we live. As we argue, the NEP captures the cognitive element of environmental orientation, and feeling connected to nature captures the affective element.

Research on connectedness to nature consistently suggests that people who feel more psychologically linked to the natural world are more likely to engage in pro-environmental behavior. Two meta-analyses found strong and consistent positive relationships between pro-environmental behavior and connection to nature [@mackayPeopleWhoFeel2019; @whitburnMetaanalysisHumanConnection2020]. In addition, a psychological connection to nature predicts environmental attitudes and behaviors beyond broader environmental worldview measures, suggesting that connectedness to nature is not just a re-labeling of general environmental concern but a distinct predictor [@nisbetNatureRelatednessScale2009]. More recent work extends this finding across multiple behavioral domains, including civic action, recycling, household conservation, and consumer choices, which were all predicted by connectedness to nature, whereas policy support and transportation were predicted by environmental concern, which included the NEP [@gkargkavouziMultidimensionalMeasureEnvironmental2019]. One study assessed the role of personality traits in environmental behavior and provided a comparison of the mediation effect of both the NEP and the CNS on environmental behaviors [@brickUnearthingGreenPersonality2016]. The personality traits of Openness and Conscientiousness were both shown to be positive predictors of environmental behavior. In both cases, the CNS had a stronger mediating effect than the NEP. In other words, the CNS better predicted pro-environmental outcomes.[^1] Finally, the CNS (plus control variables) accounted for 35% of the variance in environmental behaviors[@sparksMeasuringEnvironmentalValues2021].

[^1]: However, these authors also note that the CNS contains language that is more consistent with liberal environmentalism, and so may fail to adequately capture the full dimensionality of environmental support across the political spectrum [@brickUnearthingGreenPersonality2016]. Indeed, like the NEP, the CNS is composed mostly of items that tap into politically progressive environmentalism and thus may miss conservative pro-environmental orientation [@sparksMeasuringProenvironmentalOrientation2022].

# Conceptual Model and Hypotheses

Cultural cognition can be thought of as a deep core belief; therefore, it is an antecedent to environmental orientations. Culture-based views of risk and the social-psychological tendencies to protect those views likely precede the development of environment-specific cognition, attitudes, and orientations. Given the mixed evidence, it is unclear whether the connection between cultural cognition and environmental behavior is direct or indirect.

Taking the step from cultural worldviews to environmental orientations, we suggest that environmental orientations contain both a cognitive worldview element that can be measured by the New Ecological Paradigm (NEP) and an affective emotional connection to nature, as measured by the Connectedness to Nature Scale (CNS). As noted, cultural worldviews are an antecedent to environmental orientations. Similarly, we posit that ecological worldviews are an antecedent to connectedness to nature. Finally, environmental orientations are predictive of pro-environmental behaviors.[^2] @fig-model illustrates our conceptual model connecting cognition based on cultural and ecological worldviews, environmental orientations, including both the ecological worldview component and the affective connection to nature component, and pro-environmental behaviors.

[^2]: Our expectations and analyses are based on pro-environmental behaviors, overall. We do not distinguish between private or public behaviors.

```{r}
#| label: fig-model
#| fig-cap: "Conceptual model tested with the structural equation model: cultural cognition shapes environmental orientation—first the New Ecological Paradigm (NEP) and then the Connectedness to Nature Scale (CNS)—which in turn shapes pro-environmental behavior."
#| fig-width: 7
#| fig-height: 2.3
draw_concept_model()
```

Based on the model shown in @fig-model, we posit several hypotheses regarding the relationships between cultural cognition, the NEP, the CNS, and environmental behavior. Our hypotheses include the sequence approach shown in @fig-model, where cultural worldviews influence ecological worldviews, ecological worldviews influence connectedness to nature, and connectedness to nature influences pro-environmental behaviors. We also have direct hypotheses, where cultural and ecological worldviews are directly associated with behaviors. We expect that:

> H1: Hierarchical-individualists are associated with a *decrease* in the ecological worldview element of environmental orientation, a *decrease* in the connected to nature element of environmental orientation, and *fewer* pro-environmental behaviors

> H2: Egalitarian-communitarians are associated with an *increase* in the ecological worldview element of environmental orientation, an *increase* in the connected to nature element of environmental orientation, and *more* pro-environmental behaviors

> H3: Ecological worldviews are associated with an *increase* in connectedness to nature

> H4: Both the cognitive and affective elements of environmental orientations are associated with *more* pro-environmental behaviors

# Data and Methods

To examine the relationship between cultural cognition, environmental orientations, and behavior, we draw on a sample of 501 adults living in the US purchased through Survey Sampling Inc. Overall, the sample is largely reflective of the US population in terms of demographics, as 51% of respondents are women, 67% are white, and the average respondent is 44 years of age. Additionally, the average person in our sample had attended some college and reported a household income between \$95,001 and 130,000. The survey was administered online in July 2017. The sample size was determined based on cost considerations. Our analysis was not pre-registered.

## Measures

Our dependent variable is a 12-item scale of self-reported pro-environmental behaviors. The questions were taken from the 2000 Gallup Earth Day Poll. Respondents were given a list of several environmental behaviors and asked, “*Which of these, if any, have you, yourself, done in the past year*?” Possible responses included, “Yes, have done; No, have not done; No opinion.” Yes, responses were coded as 1, and no/no opinion was coded as 0. For the purposes of analysis, we treated the behavioral measures as observed and added the number of self-reported behaviors together to produce a measure that ranged from 0 (no behaviors) to 12 (all behaviors). @tbl-behaviors shows the question wording as well as the mean and sd for each self-reported behavior. We also note the 7 public and 5 private behaviors, but we use the single 0-12 scale for analysis.[^4]

[^4]: The results of analyses that separates public and private are substantively the same as presented here.

```{r}
#| label: tbl-behaviors
#| tbl-cap: "Self-reported pro-environmental behaviors by type (public/private), with means and standard deviations"
make_tbl1_flextable()
```

As shown in @tbl-behaviors, the mean number of self-reported pro-environmental behaviors was `r round(mean(envatt$BEHAVE12, na.rm = TRUE), 2)`, with a standard deviation of `r round(sd(envatt$BEHAVE12, na.rm = TRUE), 2)` in the 0 to 12 scale. For public behaviors the mean was `r round(mean(envatt$public, na.rm = TRUE), 2)`, with a standard deviation of `r round(sd(envatt$public, na.rm = TRUE), 2)`, and the mean number of self-reported private behaviors was `r round(mean(envatt$private, na.rm = TRUE), 2)`, with a standard deviation of `r round(sd(envatt$private, na.rm = TRUE), 2)`. Overall, respondents reported far more private than public behaviors. In terms of the behavior themselves, recycling had the highest mean at `r round(mean(envatt$recycle, na.rm = TRUE), 2)`, which indicates that `r round(mean(envatt$recycle, na.rm = TRUE), 2) * 100`% of the respondents reported that they recycle, followed closely by reducing household energy use at `r round(mean(envatt$energy, na.rm = TRUE), 2)`. Buying a product because it was better for the environment had the lowest mean among private behaviors at `r round(mean(envatt$buycott, na.rm = TRUE), 2)`. Among public behaviors, voting for or working for a candidate because of their environmental positions had the highest mean at `r round(mean(envatt$cand, na.rm = TRUE), 2)`, followed by signing a petition supporting environmental protection at `r round(mean(envatt$petition, na.rm = TRUE), 2)`. Contacting a business to complain about environmentally harmful products/policies had the lowest mean among public behaviors at `r round(mean(envatt$cont.bus, na.rm = TRUE), 2)`.

Our independent variables are composite measures of cultural worldviews, ecological worldviews (New Ecological Paradigm), and the Connectedness to Nature Scale. For cultural cognition, we rely on the short form of the cultural cognition survey questions, which include three questions for each of the four cultural worldviews [see @kahanCulturalCognitionConception2012]. Respondents were asked to indicate their agreement on a 1 (strongly disagree) to 4 (strongly agree) scale with three statements for each cultural worldview. To measure the New Ecological Paradigm (NEP) we used agreement with three statements on a 1 to 5 scale (1 - strongly disagree, 3 - neutral, 5 - strongly agree). Finally, for the Connectedness to Nature Scale (CNS), respondents were asked to indicate their agreement on a 1 to 5 scale (1 - strongly disagree, 3 - neutral, 5 - strongly agree) with three statements. For both the NEP and CNS, a higher value indicates a stronger orientation towards the environment.

@tbl-measurement shows the question wording for each measurement item used for the latent constructs as well as the fit statistics for each of the constructs including Cronbach’s $\alpha$, Composite Reliability (CR), and the Average Variance Extracted (AVE). As shown in @tbl-measurement, each construct showed sufficient reliability (α and CR \>0.70), except for the communitarianism scale. However, CR values between 0.60 and 0.70 are acceptable for exploratory analysis [@hairPrimerPartialLeast2015].

```{r}
#| label: tbl-measurement
#| tbl-cap: "Measurement items and construct reliability (Cronbach's α, composite reliability, average variance extracted)"
make_tbl2_flextable()
```

For the cultural worldview measures, we combined the scales to create dummy variables for hierarchical individualists and egalitarian communitarians. Respondents who scored at or above the median for both scales were coded as 1, and all others were coded as 0 [similar to @kahanPolarizingImpactScience2012]. Overall, `r sum(envatt$HIER_INDIV == 1, na.rm = TRUE)` respondents (`r round(mean(envatt$HIER_INDIV == 1, na.rm = TRUE) * 100, 1)`%) were classified as hierarchical individualists and `r sum(envatt$EGAL_COMM == 1, na.rm = TRUE)` respondents (`r round(mean(envatt$EGAL_COMM == 1, na.rm = TRUE) * 100, 1)`%) as egalitarian communitarians.

To test our hypothesis, we estimated a Structural Equation Model (SEM) using the `lavaan` package in `R` [@rosseelLavaanPackageStructural2012]. The dependent variable, self-reported pro-environmental behaviors, is an observed variable, and the cultural cognition worldviews, the CNS, and the NEP scale are latent constructs using the measures in @tbl-measurement. In addition, we examine the mediation effects of environmental orientation by examining the direct, indirect, and total effects of cultural worldviews on pro-environmental behavior.[^5]

[^5]: In the analysis we control for demographic variables, including age, gender (male=1), race (white=1), education, and income. We chose white males as the indicator variables because of the well-known *white-male effect* associated with risk perceptions and cultural cognition [e.g., @kahanCultureIdentityProtectiveCognition2007]. 

# Results

The SEM results are shown in @fig-structural. For clarity, we only include the structural model and only the paths that were statistically significant at p\<0.05.

::: {#fig-structural}
```{r}
#| fig-width: 9
#| fig-height: 7.5
draw_sem_struct_hiec(fitFullHIEC)
```

Structural model of cultural worldviews, ecological worldviews, connectedness to nature, and pro-environmental behaviors.
:::

As shown in @fig-structural, as expected, hierarchical individualists were negatively associated with ecological worldviews (NEP). This was expected because a positive value indicates a more environmentally centric worldview. Also as expected, egalitarian-communitarians were positively associated with the NEP. Additionally, the cognitive element of environmental orientations, the NEP, was positively associated with the affective element, the CNS, as expected. Regarding pro-environmental behaviors, as expected, both elements of environmental orientation were positively associated with an increase in the number of behaviors; however, there was no direct relationship between cultural worldviews and behavior. Finally, egalitarian-communitarians were more likely to feel connected to nature; therefore, they had the strongest environmental orientation, as both the cognitive and affective elements were significant.

The path coefficients are standardized, which means that they equal the expected change in the standard deviation (sd) of the outcome variables for each standard deviation of the predictor, keeping all other variables constant. For example, the path coefficient from CNS to behaviors is 0.404; therefore, holding the other variables constant, a 1-sd increase in the CNS scale is associated with a 0.404-sd increase in the number of behaviors. In terms of the number of actual predicted behaviors, as shown in @tbl-behaviors, the sd of behaviors is 2.86, and 0.404 × 2.86 equals 1.155 predicted additional behaviors for each 1-sd increase in the CNS.

Standardized path coefficients allow for a comparison of the strengths of the various relationships. As shown, the strongest relationship is between the two environmental orientation measures, with the NEP to CNS path = 0.425. In terms of the relationship to behaviors, the CNS path was the strongest at 0.404, followed by the NEP to behavior path at 0.240. No other paths to behaviors were significant. For cultural worldviews to environmental orientations, hierarchal individualists have the strongest (negative) relationship, at -0.392, with ecological worldviews. Egalitarian communitarians has a stronger relationship with ecological worldviews (0.231) than with connectedness to nature (0.188), but is significantly associated with both.

The fit statistics for the model are shown in @tbl-fit-full and include the Root Mean Square Error of Approximation (RMSEA), Comparative Fit Index (CFI), Tucker-Lewis Index (TFI), Root Mean Square Error of Approximation (RMSEA) and its confidence intervals, Standardized Root Mean Square Residual (SRMR), and AIC and BIC. Overall, the statistics indicate a reasonable fit for the model [@huCutoffCriteriaFit1999].

```{r}
#| label: tbl-fit-full
#| tbl-cap: "Model fit statistics for the structural model"
make_tbl_fit_full_flextable()
```

Next, we examined the direct, indirect, and total effects of cultural worldviews on behavior. Those results are shown in @tbl-serial-effects.

```{r}
#| label: tbl-serial-effects
#| tbl-cap: "Standardized decomposition of cultural worldview group effects on pro-environmental behavior."
make_tbl_serial_effects_flextable()
```

For hierarchical individualists, the total effect on behaviors was -0.121, and for egalitarian communitarians, it was 0.199. However, as shown in @fig-structural and @tbl-serial-effects, there is no direct relationship between cultural worldviews and behaviors; therefore, the effects are strictly a result of indirect effects, as environmental orientations appear to mediate the relationship between cultural worldviews and behaviors.

The serial effect includes the paths of cultural worldviews to the NEP, NEP to CNS, and CNS to behaviors. For example, for hierarchical individualists, the serial effect is calculated as −0.392 × 0.425 × 0.404 = −0.067. The total indirect effect captures the paths from cultural worldviews to NEP, NEP to behavior, cultural worldviews to CNS, and the serial effect. For hierarchical individualists that is (−0.392×0.240) + (0.010×0.404) + (−0.067) = −0.094 + 0.004 − 0.067 = −0.157. Finally, the total effect is simply the direct effect plus the total indirect effect.

# Discussion

Scholars have noted a gap between values and beliefs about the environment and pro-environmental behavior. In particular, cultural worldviews derived from cultural cognition, specifically egalitarianism and communitarianism, have been shown to inform views on several environmental issues as well as overall attitudes toward the environment. However, little research has found a connection between cultural worldviews and environmental behavior [but see @markleUnderstandingProEnvironmentalBehavior2019]. In this paper, we developed and tested a conceptual model based on the existing VBN and ACF frameworks, in which cultural worldviews influence environmental orientations, and environmental orientations influence behaviors. Additionally, our contribution notes that cultural and ecological worldviews are cognitive features of belief systems that provide both a basis of values and beliefs as well as a series of psychological mechanisms that ensure that beliefs and attitudes are consistent with the underlying worldviews. In turn, these cognitive features of belief systems influence the affective element of environmental orientations, which is the emotional connection to nature. Finally, the affective element of environmental orientation is most strongly associated with pro-environmental behavior.

Based on our conceptual model, we posited that egalitarian communitarians would be positively associated with environmental orientations and environmental behavior, and hierarchical individualists would be negatively associated with environmental orientations and environmental behavior. Additionally, we expected that the cognitive element of environmental orientation would be associated with the affective element and that both elements of environmental orientation would be associated with increased self-reported pro-environmental behaviors. Overall, we found mixed results for our hypotheses. As expected, egalitarian communitarians were associated with both the CNS and NEP; however, they were not directly associated with behavior. In addition, hierarchical individualists were less likely to have an ecological worldview, but they were no more or less likely to feel connected to nature than those with other cultural worldviews. Finally, both elements of environmental orientation were associated with more pro-environmental behaviors, and based on the standardized coefficients, the affective element (CNS) had a stronger relationship than the cognitive element (NEP).

Given the uncertainty about the relationship between cultural worldviews and pro-environmental behaviors, we also conducted a mediation analysis to examine whether environmental orientations mediate the relationship between cultural worldviews and behaviors. If so, cultural worldviews would still influence behavior but only indirectly. We found that both the cognitive and affective elements of environmental orientations mediated the relationship between cultural worldviews and environmental behaviors. This means that the effect of cultural worldviews on pro-environmental behaviors is based solely on their influence on environmental orientations. Along with the results of the structural equation modeling, this helps us understand cultural cognition as an antecedent of environmental orientation and fill some of the attitude-behavior gap.

We also found that the Connectedness to Nature Scale (CNS) had more predictive power regarding environmental behaviors than the NEP, based on the standardized path coefficients shown in @fig-structural. This finding is in line with work showing that the CNS is a stronger predictor than the NEP [e.g., @sparksMeasuringEnvironmentalValues2021]. Additionally, the CNS and NEP measure two distinct concepts that can be grouped under the umbrella of environmental orientation, where the NEP is explicitly a measure of environmental attitude, while the CNS is a measure of emotional connection to nature. Overall, the affective element of environmental orientations is more proximate to behaviors than the cognitive element. Similarly, a recent study found that harnessing the right mix of emotions is a powerful motivator of climate action [@kovacsActingWeFeel2024].

Another reason that the affective element of environmental orientation, the CNS, displays more power than the cognitive NEP to influence behavior is that there has been a slow decline in environmental concern as measured by the NEP [@louMetaanalysisTemporalShifts2022]. The downtrend of the NEP was shown through meta-analysis; however, there was no indication of why this may be happening. We are not aware of a similar study tracking CNS over time, but that may prove to be a fruitful avenue for further study. If only the NEP is declining over time, that may indicate that the measurement instrument is becoming outdated, potentially because of shifts over time making the distinction between the dominant social paradigm and the new ecological paradigm less apt. For example, many see technological innovation as the key to unlocking a new green economy rather than destroying the earth. Relatedly, perhaps the language used in the NEP has just become outmoded and out of line with how environmentalists currently would express environmental attitude [see @sparksMeasuringProenvironmentalOrientation2022 for a discussion of this].

There are a few limitations with this paper. First, the sample size was relatively small (n = 501). While we have no reason to think that a larger sample size would substantively change the results, the results should be replicated with more respondents. Additionally, our data are cross-sectional, and future research should measure the nature of the relationships between cultural cognition, environmental orientation, and environmental behavior over time using panel data. Panel data could explore the strength and consistency of the relationships over time. Finally, causality is difficult to determine using SEM with cross-sectional data. While we argue that there are theoretical and empirical reasons to see cognition preceding the affective elements of belief systems, we recognize that the statistical associations we found do not constitute definitive evidence of causality. Therefore, we have been careful in our language when describing our findings. 


# Conclusion

<!-- cognitive precedes affective because of the empirical proximity to behaviors; sequence based on previous research -->

Achieving environmentally sustainable outcomes requires significant engagement in pro-environmental behavior across all segments of society. While a large body of research exists that examines pro-environmental behaviors, some gaps remain. In this paper, we developed a conceptual model that draws a distinction between the cognitive and affective elements of belief systems and tested hypotheses based on that approach. We argue that combining cultural cognition and environmental orientation should improve our understanding of the drivers of environmentally protective behaviors 

Because cultural cognition shapes how an individual responds to risk [e.g., @kahanPolarizingImpactScience2012], it was something of a puzzle as to why it only weakly correlates to behavior that protects the environment – an action that can be seen as a way to reduce environmental risk to the individual. The evidence presented in this study suggests that environmental orientation, whether attitude (NEP) or emotional connection (CNS), is an important link in this relationship. Additionally, we presented evidence that cultural worldviews are best thought of as indirect predictors of pro-environmental behavior. This relationship was mediated by environmental orientation, as measured by both the NEP and CNS. Furthermore, by explaining behavior with a mediation model, we could close some of the so-called attitude-behavior gap. When cultural cognition and environmental orientation are aligned, people are more likely to engage in environmentally protective behaviors.

Another implication of this study is further evidence that CNS is a stronger predictor of behavior than NEP. This may be useful for researchers trying to decide which measures of pro-environmental orientation to include in their questionnaires. We would like to see further testing of this finding in studies in other contexts and through meta-analytic techniques.

```{=html}
<!-- Altogether, . 

Specifically, in this paper, we suggest that environmental orientation mediates the relationship between cultural cognition and environmental behavior. Essentially, environmental orientation helps connect cultural worldviews and risk perceptions to action to reduce that risk. Moreover, cultural cognition helps to bridge the attitude-behavior gap by providing a basis through environmental orientations derive and subsequently influence behavior.



In addition, environmental orientation is likely associated with the myths of nature connected to the various cultural worldviews. As a result, environmental orientation, as measured by NEP and/or CNS, likely mediates the relationship between cultural cognition and environmental behavior. Figure 1 


To further elaborate the attitude-behavior model we examine the relationship between cultural cognition, environmental orientation (attitude and emotional connection), and behavior.


In this paper, we examine whether environmental orientations mediate the relationship between cultural cognition and behavior such that cultural cognition shapes environmental orientations, which in turn influences pro-environmental behavior. In doing so, we argue that when cultural cognition aligns with environmental orientation, there is a stronger relationship with pro-environmental behaviors than when considering attitude or affect alone, thus reducing the attitude-behavior gap.

-->
```

\newpage

# References